---
title: "Agriculture Machine Learning (Classification Prediction)"
author: "Kar"
date: "2021"
output: 
  github_document:
    toc: true
    toc_depth: 3
always_allow_html: yes

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1 R PACKAGES

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(skimr)
library(kableExtra)
library(corrplot)
library(caret)
library(MASS)

```

# 2 INTRODUCTION



# 3 DATA IMPORT AND CLEANING

This project uses a public crop dataset in kaggle.com [Link](https://www.kaggle.com/atharvaingle/crop-recommendation-dataset), named "Crop Recommendation Dataset" by Artharva Ingle. 

## 3.1 Data import

The data set is uploaded into R using following code, and the table indicates a successful import. 

```{r}
crop <- read.csv("Crop_recom.csv",
                 fileEncoding = "UTF-8-BOM")

```

The table shows soil nutrients, N, P, and K, as well as temperature, humidity, pH, rainfall, and label. There are no unit for N, P, and K and the values are in term of N-P-K ratio in the soil.

The dataset has following descriptions, adapted from the Kaggle website. 

```{r}
variables <- c("N", "P", "K", "temperature", "humidity", "ph", "rainfall", "label")

description <- c("N-P-K Ratio of Nitrogen (N) portion in soil",
                 "N-P-K Ratio of Phosphorus (P) portion in soil",
                 "N-P-K Ratio of Potassium (K) portion in soil",
                 "temperature in degree Celsius, oC",
                 "relative humidity in %",
                 "Soil pH value",
                 "rainfall in mm",
                 "Different crops")

data.frame(variables, description) %>% 
  kbl() %>% 
  kable_styling(bootstrap_options = c("hover", "stripped", "bordered"))

```

## 3.2 Data exploration

This data has 2200 rows of observations and 8 columns of variables. 

* The "label" is labelled as a character variable.

* And the rest are numerical variables.

* The data set has obviously been cleaned and having no missing values that needs to be managed by looking at the **complete_rate** and the associated column **n_missing** that used to detect missing values in the data frame.

```{r}
skim_without_charts(crop)
```

Numerical variables either labelled as "int", "num", or "dbl" in R, which the last "dbl" is usually used to labeled floating numbers. 

```{r}
str(crop)

```

## 3.3 Data manipulation

**Convert ”label“ to factor**

To quickly examine what are the levels within the variable "label", which is important as it tells the type of crops within this dataset, I will convert the variable "label" from character into factor. It will help examination using R functions and later analysis.

```{r}
# To protect the original dataset I will hereby create a new object named "crop2" to carry the cleaned data set from this project.

crop2 <- crop %>% 
  mutate(label = as.factor(label))

```

Following result shows that there are 22 crops in this data set. Each crop has 100 samples, which is excellent as sample sizes are equal and will make statistical comparison fairer. 

```{r}
summary(crop2$label)

```

**Reduce decimal places**

This step is optional but I decided to make it happened. I will restrict the decimal places of "temperature", "humidity", "pH", and "rainfall" to only one, as this length of decimal places is sufficient.

```{r}
crop2 <- crop2 %>% 
  mutate(temperature = round(temperature, 1),
         humidity = round(humidity, 1),
         ph = round(ph, 1),
         rainfall = round(rainfall, 1))
  
```

Following shows statistics of all the columns within the dataset such as minimum, maximum, median, and mean. It also shows the number of samples if the column is in factor format, such as the "label" that I have converted it from character format.

```{r}
summary(crop2)

```


# 4 Exploratory Data Analysis

## 4.1 Histograms

It will be interesting to visualise the distribution of each numerical variables in the data set as a initial examination.

```{r}
# set up data frame for this section

df4.1 <- crop2 %>% 
  pivot_longer(c(1:7),
               values_to = "result",
               names_to = "variables") %>% 
  mutate(variables = factor(variables, levels = c("N",
                                                     "P",
                                                     "K",
                                                     "temperature",
                                                     "humidity",
                                                     "ph",
                                                     "rainfall"))) 



# plot

ggplot(df4.1, aes(x = result, fill = variables)) +
  geom_histogram(colour = "white") +
  facet_wrap(~variables, scales = "free_x") +
  labs(x = "Variables",
       title = "Distribution of Numerical Variables") +
  theme_bw() +
  theme(legend.position = "none",
        plot.title = element_text(face = "bold")) 

```
Insights: 

* Distributions of N, P, and K are quite wide spread which might be due to the presence of different soil types contributing to this dataset. It indicating that they can be counding variables in this dataset.

* The majority of temperatures are spread between 20 - 30oC in this dataset.

* Humidity is quite wide spread with majority fall between 75 - 100%.

* pH and in soil are most slightly acidic with a value around 6.

* Rainfalls in this entire dataset are less than 300mm, with majority fall between 50 - 120mm. 


## 4.2 Correlogram 

Following correlogram (a plot for correlation) shows that variables are independent from each other except P and K. 

There is a strong relationship between P and K with a correlation of 0.74. It might be an issue during model building however VIF will be checked again in that section to avoid multicollinearity. Though a correlation less than 0.8 between two variables should be indicating that multicollinearity might not exist, but VIF will be computed.  

```{r}
# convert into matrix, remove the factor "label" 

cor_c <- cor(crop2[, 1:7])

# correlogram

corrplot(cor_c, method = "number", type = "upper")

```

## 4.3 Scatter plot

Plotting the relationship between P and K. 

```{r}
ggplot(crop2, aes(x = P, y = K, colour = label)) +
  geom_point(alpha = 0.5) +
  theme_bw() +
  labs(title = "Positive Relationship betweem P and K") 

```
From the graph, there might be a positive relationship between K and P, however the high value of correlation 0.74 might be due to the type to crops and especially the crop groups near 150.

## 4.4 Boxplots

This section uses boxplot to simultaneously compare the 7 numerical variables among different crops stored within the vector, "label". Recall, the 7 numerical variables are:

```{r}
names(crop2[1:7])

```


```{r, fig.width = 8, fig.height=20}
# set up data frame

df4.3 <- crop2 %>% 
  pivot_longer(c(1:7),
               values_to = "result",
               names_to = "variables") %>% 
  mutate(variables = factor(variables, levels = c("N",
                                                     "P",
                                                     "K",
                                                     "temperature",
                                                     "humidity",
                                                     "ph",
                                                     "rainfall"))) 

# plot boxplots

ggplot(df4.3, aes(x = label, y = result, colour = label)) +
  geom_boxplot() +
  facet_wrap(~variables, scale = "free", ncol =1) +
  theme_bw() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.2))
  
```


# 5 Model Building 

The purpose of modeling of this project is to build an effective predictive model to suggest the most appropriate crops to grow in a particular farm based on various parameter. 

## 5.1 Variables Selection

This section aims to select relevant and confounding variables during model building. According to the purpose of this project all of these variables are agricultural-wise important and relevant in determining the type of crops to grow in a particular farm.

```{r}
head(crop2)

```

Therefore, I will keep all of these variables.


## 5.2 Train-Test Split

The column "label" is what we are going to predict. 

```{r}
set.seed(123)

# Create Data Partition

training.set <- crop2$label %>% createDataPartition(p = 0.8, list = F)

# split 80% for train set and 20% for test set

train.data <- crop2[training.set, ]

test.data <- crop2[-training.set, ]

```

## 5.3 Discriminant Analysis  

5 methods of Discriminant Analysis

```{r}
# Create data parameter 

preprocess_parameter <- train.data %>% preProcess(method = c("center", "scale"))

# Data standardisation

train.transformed <- preprocess_parameter %>% predict(train.data)
test.transformed <- preprocess_parameter %>% predict(test.data)

```

**LDA**

```{r}


```






* train validate test split (train and validate, then use test test to makesure effective in real world)
* multiple linear regression
* Lasso regression (because this dataset is going to be so sparse with all these dummy variable, lasso help to normalise that)
* Random forest (so we have a tree-based model to compare to our LM)
  * Can use gradient boosted tree
  * Support vector regression (classification)
* tune models GridsearchCV
* test ensembles







# 6 Model for Production




# Conclusion




